Model Recognition as Planning

### 2018-06-20 doc/aaai-19/model-recognition.tex
added goals to the definition of the model recognition task

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The {\em plan length} is denoted with $|\pi|=n$. A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.

@@

\section{Model Recognition}
\label{sec:recognition}
Given $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of planning models and $\tau$, the observation of a plan execution, {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.

Our approach is to estimate this probability according to the amount of {\em edition} required by the action schemas in $\mathcal{M}$ to produce a plan that is compliant with $\tau$. We say that a plan $\pi$ is compliant with the observation of a plan execution $\tau$ when (1) the observed actions $\pi$ but with certain actions omitted, (2) the observed states is the sequence of states traversed by $\pi$ with certain states omitted or cerntain fluents in the states omitted.

To estimate the probability of a model in $M$ to produce an observation $\tau$ of a plan execution we assume that:
\begin{enumerate}
\item All the planning models in $M$ are {\em comparable} \strips\ models.
\item $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{m}}$ is an interleaved sequence of {\em partially-observed} states and actions. The conjunction of these two assumptions makes that transitions between two consecutive observed states may involve the execution of more than a single action. Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound. This means that the actual length of the plan whose execution is observed is unknown. 
\end{enumerate}

@@

\section{Model Recognition as Planning}
\label{sec:asPlanning}
For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation. The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. 
\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\tau$}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\tau$.
\end{enumerate}

>>>>

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that, when executed starting from the initial state $I$, induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The {\em plan length} is denoted with $|\pi|=n$. A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.

\subsection{Observations of plan executions}
An observation $\tau$ of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combinations of a sequence of {\em partially-observed} actions $\tup{a_1,\ldots,a_n}$ and a sequence of {\em partially-observed} states $\tup{s_1,\ldots,s_m}$ such that $n\leq |\pi|$ and $m\leq |\pi|+1$. Note that, because of the partial observability, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action. Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound. This means that the actual length of the plan whose execution is observed is unknown.

The observation of a plan execution $\tau$ is {\em compliant} with a plan $\pi$ iff:
\begin{itemize}
\item The sequence of actions in $\tau$ is the same sequence of actions $\pi$ but with certain actions omitted.
\item The sequence of states in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states.
\end{itemize}

The observation of a plan execution $\tau$ is {\em noiseless} iff the value of any fluent or any action observed in $\tau$ is correct.

@@

\section{Model Recognition}
\label{sec:recognition}
Given a classical planning problem $P=\tup{F,A,I,G}$ where the semantic of the actions in $A$ is unknown. $\tau$ the observation of the execution of a plan for solving $P$ and $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of action models that define different semantics for the actions in $A$, then {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.

Our approach is to estimate this probability, for each $\mathcal{M}\in M$, according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan that is compliant with $\tau$.

@@

\section{Model Recognition as Planning}
\label{sec:asPlanning}
For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation that follows these assumptions:
\begin{enumerate}
\item Observations in $\tau$ are {\em noiseless}. 
\item The initial state $I$ is {\em fully observed}. This means that the set of fluents $F$, and the corresponding set of predicates $\Psi$ that shapes the fluents in $F$, are inferrable from $I$.
\end{enumerate}

The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. 
\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\tau$}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\tau$.
\end{enumerate}
